The \project{unRAVE} catalog

### 2016-09-07 article/unrave.tex
Hot stars #24; abundance limits #13; sub-giant artefact #24

The fibre fringing is another example of how test spectra that are
unusual from the training set can be projected as peculiar artefacts in the
label space.  In other words, unusual spectra can appear as `clumps' in
regions of parameter space that we could consider as being normal 
(e.g., Solar-type stars).  We addressed this issue for the main-sequence
and giant models by using a joint model to calculate relative weights.
However, spectra that are unusual from the joint model training set can
still project as systematic artefacts.  An interrogation of our results
has identified one such group of metal-poor ($[{\rm Fe/H}] \approx -0.4$)
stars near $\teff \approx 3250$~K, which peculiarly sit above the isochrone. 
The \rave\ morphological classifications \citep{Matijevic_2012} indicate 
these stars have chromospheric emission, a feature not present in our 
training set.  Based on this (and other previously discussed) examples,
we recommend either a thorough investigation of unexpected structure\footnote{
Before submitting to \emph{Nature}} 
using all the available 
information from \rave, or judicious quality controls (e.g., requiring the 
morphological classification to be `normal') to remove these spurious projections.

\stub{Sub-giant artefact; comment as projection}

\stub{Hot stars!}

\stub{based on our interrogation of the data, we supply a 'QC' column that uses very strict requirements to eliminate weird things: morphological classifications, SNR, R (pipeline), $\chi_r^2$, etc....}

>>>>

The fibre fringing is another example of how test spectra that are
unusual from the training set can be projected as peculiar artefacts in the
label space.  In other words, unusual spectra can appear as `clumps' in
regions of parameter space that we could consider as being normal 
(e.g., Solar-type stars).  We addressed this issue for the main-sequence
and giant models by using a joint model to calculate relative weights.
However, spectra that are unusual from the joint model training set can
still project as systematic artefacts.

Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
first is a small over-density at the base of the giant branch, which is
likely a consequence of joining the 9-label and 3-label models.  The
second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
present in our training set, yet there are many in the \rave\ survey.
However, the \rave\ pre-processing pipeline (\texttt{SPARV}) performs
template matching against grids of cool \emph{and} hot stars, and therefore
we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
where each hexagonal bin is colored by the \emph{maximum} temperature
reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
at just above the turn-off point in our results.

\stub{this following paragraph needs rewording}
Based on a thorough interrogation of our results, we have adopted two boolean
columns to signify whether our labels meet specific quality constraints:
one marked \texttt{QC} to indicate reliable stellar parameters ($\teff$, $\logg$, [Fe/H]), and \texttt{QC\_ABUNDANCE} to indicate reliable chemical abundances.
For stellar atmospheric parameters to meet our quality constraints, we
require: $\chi_r^2 < 3$; $S/N >10$~pixel$^{-1}$; and \texttt{TeffS} $< 8000$~K
in order to remove hot stars that are outside the training set.  We
specified the quality constraint in chemical abundances to further
require that the iron abundance $[{\rm Fe/H}] > -1.5$, since this approximately
corresponds to the point where individual abundances will become undetectable
in \rave\ spectra, and the individual abundance labels will be driven by our
constraint to place them on the mean galactic chemical evolution trend.
These quality constraints should not be taken verbatim: judicious use of
the morphological classifications \citep{Matijevic_2012} or metadata from
the pre-processing pipelines is encouraged.
